\section{HDR Floating-Point Representation}
\label{sec:hdr-float}

Deep-zoom Mandelbrot rendering places unusual demands on numerical
representation. Orbit values may span an extreme dynamic range over the course
of iteration, while small relative differences between nearby orbits must be
tracked accurately. Standard floating-point formats are poorly matched to this
combination of requirements: fixed-precision mantissas limit relative accuracy
at large magnitudes, while arbitrary-precision formats incur prohibitive
computational cost when used per pixel.

To address this gap, we employ a custom \emph{high dynamic range (HDR)}
floating-point representation that explicitly separates scale and precision
(\cref{fig:type-hierarchy}).
This representation is designed to preserve relative accuracy across a wide
range of magnitudes while remaining efficient on GPUs.

\subsection{Decoupling scale and precision}

In standard IEEE floating-point arithmetic, each value is represented as a
mantissa multiplied by a power of two, with both components stored implicitly in
a single word. While convenient, this tightly couples scale and precision: as
values grow in magnitude, fewer bits remain available to represent small
relative differences.

The HDR representation instead stores values in the form
\begin{equation}
x = m \cdot 2^{e},
\end{equation}
where:
\begin{itemize}
\item \(m\) is a fixed-precision mantissa stored in standard floating-point or
      double-double format, and
\item \(e\) is an explicitly managed integer exponent.
\end{itemize}

By storing the exponent separately, the mantissa can remain centered near unit
magnitude regardless of the overall scale of the value. This preserves relative
precision even when the absolute value becomes extremely large or small.

\subsection{Normalization and arithmetic}

HDR arithmetic maintains the invariant that the mantissa remains within a fixed
normalized range. After each arithmetic operation, the mantissa is renormalized
and any excess scaling is folded into the exponent. Conceptually, this mirrors
the behavior of floating-point normalization, but with explicit control over the
process.

Arithmetic operations are performed as follows:
\begin{itemize}
\item \textbf{Multiplication} combines mantissas and adds exponents.
\item \textbf{Addition and subtraction} align exponents explicitly before
      operating on mantissas.
\item \textbf{Renormalization} ensures that mantissas remain well-scaled and
      numerically stable.
\end{itemize}

Because mantissas are stored using fixed-precision types, these operations map
naturally to GPU hardware while avoiding the overhead of general-purpose
arbitrary-precision arithmetic.

\subsection{HDR complex numbers}

Mandelbrot iteration operates on complex numbers. In the HDR formulation, each
complex value
\[
z = x + i y
\]
is represented by storing separate HDR values for the real and imaginary
components. Both components share the same conceptual structure but may have
independent exponents.

This representation allows the magnitude of \(z\) to grow or shrink over many
orders of magnitude while maintaining accurate relative phase and magnitude
information, which is critical for both escape testing and perturbation-based
methods.

\subsection{Role of HDR in perturbation and reference orbits}

The HDR representation plays a complementary role to perturbation (\cref{sec:perturbation-concept}). In the
perturbation framework, the reference orbit (\cref{sec:ref-orbit-calc}) typically grows rapidly in magnitude
and therefore demands high dynamic range, while perturbation deltas remain small
and can often be represented using lower precision.

HDR arithmetic is used primarily for:
\begin{itemize}
\item computing reference orbits at deep zoom levels,
\item evaluating linear and nonlinear terms involving large reference values,
\item maintaining numerical stability across long iteration sequences.
\end{itemize}

By decoupling scale from precision, HDR arithmetic allows the reference orbit to
be computed efficiently without sacrificing accuracy, while still enabling the
mixed-precision structure exploited by perturbation rendering.

\subsection{Advantages over conventional representations}

Compared to standard floating-point formats, HDR arithmetic offers:
\begin{itemize}
\item substantially increased dynamic range,
\item stable relative precision independent of magnitude,
\item predictable and controllable numerical behavior.
\end{itemize}

Compared to arbitrary-precision libraries, it avoids dynamic memory allocation,
irregular control flow, and large per-operation overheads, making it well suited
for parallel execution on GPUs.

\subsection{Summary}

The HDR floating-point representation provides a numerically robust and
computationally efficient foundation for deep-zoom Mandelbrot rendering. By
explicitly separating scale and precision, it bridges the gap between standard
floating-point arithmetic and full arbitrary-precision methods. In combination
with perturbation, it enables accurate, high-performance rendering across
extreme zoom levels while remaining compatible with GPU execution models.

\begin{figure}[!htbp]
\centering
\begin{tikzpicture}[
    >=Stealth,
    typebox/.style={draw, rounded corners=2pt, minimum height=0.55cm,
                    minimum width=1.8cm, font=\scriptsize\ttfamily,
                    fill=blue!8, align=center},
    bitlabel/.style={font=\tiny, gray},
]

% ═══ Part (a): precision hierarchy ═══
\node[font=\scriptsize\bfseries] at (1.6, 5.8) {(a) Precision hierarchy};

% Row 0: native types
\node[typebox] (float) at (0, 0) {float};
\node[bitlabel, right=1pt] at (float.east) {24\,b};
\node[typebox] (double) at (3.2, 0) {double};
\node[bitlabel, right=1pt] at (double.east) {53\,b};

% Row 1: 2-component
\node[typebox] (dblflt) at (0, 1.4) {dblflt};
\node[bitlabel, right=1pt] at (dblflt.east) {${\sim}$48\,b};
\node[typebox] (dbldbl) at (3.2, 1.4) {dbldbl};
\node[bitlabel, right=1pt] at (dbldbl.east) {${\sim}$104\,b};

% Row 2: 4-component
\node[typebox] (gqf) at (0, 2.8) {GQF};
\node[bitlabel, right=1pt] at (gqf.east) {${\sim}$96\,b};
\node[typebox] (gqd) at (3.2, 2.8) {GQD};
\node[bitlabel, right=1pt] at (gqd.east) {${\sim}$212\,b};

% Row 3: arbitrary precision
\node[typebox, fill=gray!12, minimum width=3.0cm] (mpir) at (1.6, 4.2)
  {MPIR / HpSharkFloat};
\node[bitlabel, right=1pt] at (mpir.east) {arbitrary};

% Arrows: increasing precision
\draw[->, blue!40, thick] (float) -- (dblflt);
\draw[->, blue!40, thick] (double) -- (dbldbl);
\draw[->, blue!40, thick] (dblflt) -- (gqf);
\draw[->, blue!40, thick] (dbldbl) -- (gqd);
\draw[->, blue!40, thick] (gqf.north) -- ([xshift=-6pt]mpir.south west);
\draw[->, blue!40, thick] (gqd.north) -- ([xshift=6pt]mpir.south east);

% Column labels
\node[font=\tiny, gray, rotate=90, anchor=south] at (-1.3, 1.4)
  {fp32-based};
\node[font=\tiny, gray, rotate=90, anchor=south] at (4.7, 1.4)
  {fp64-based};

% HDRFloat<T> wrapper
\begin{scope}[on background layer]
  \node[draw, dashed, rounded corners=6pt, orange!70!black, thick,
        inner xsep=10pt, inner ysep=8pt,
        fit=(float)(double)(gqf)(gqd)] (hdrbox) {};
  \node[font=\scriptsize, orange!70!black, anchor=south east]
    at (hdrbox.north east) {\code{HDRFloat<T>} wrapper};
\end{scope}

% Precision arrow
\draw[<-, thick, gray!50] (-1.8, -0.3) -- (-1.8, 4.5);
\node[font=\tiny, gray, rotate=90, anchor=south] at (-2.1, 2.1)
  {increasing precision};

% ═══ Part (b): HDR representation ═══
\begin{scope}[shift={(7.5, 0.5)}]
  \node[font=\scriptsize\bfseries] at (2.0, 5.3)
    {(b) HDR representation};

  % IEEE 754 double
  \node[font=\scriptsize, anchor=east] at (-0.1, 3.65)
    {IEEE 754 \code{double}:};

  \draw[fill=gray!20] (0, 3.3) rectangle (0.4, 4.0);
  \node[font=\tiny, rotate=90] at (0.2, 3.65) {S};
  \draw[fill=blue!15] (0.4, 3.3) rectangle (1.5, 4.0);
  \node[font=\tiny] at (0.95, 3.65) {exp};
  \draw[fill=blue!8] (1.5, 3.3) rectangle (4.0, 4.0);
  \node[font=\tiny] at (2.75, 3.65) {mantissa};

  \draw[decorate, decoration={brace, amplitude=4pt, mirror}, thick]
    (0, 3.15) -- (4.0, 3.15)
    node[midway, below=5pt, font=\tiny]
      {64 bits, tightly coupled};

  % HDRFloat<T>
  \node[font=\scriptsize, anchor=east] at (-0.1, 1.45)
    {\code{HDRFloat<T>}:};

  \draw[fill=orange!20, draw=orange!70!black, thick]
    (0, 1.1) rectangle (1.5, 1.8);
  \node[font=\tiny] at (0.75, 1.45) {int32 exp};

  \node[font=\scriptsize] at (1.75, 1.45) {$+$};

  \draw[fill=blue!12, draw=blue!60!black, thick]
    (2.0, 1.1) rectangle (4.0, 1.8);
  \node[font=\tiny] at (3.0, 1.45) {\code{T} mantissa};

  \draw[decorate, decoration={brace, amplitude=4pt, mirror},
        thick, orange!70!black]
    (0, 0.95) -- (1.5, 0.95)
    node[midway, below=5pt, font=\tiny, orange!70!black] {range};
  \draw[decorate, decoration={brace, amplitude=4pt, mirror},
        thick, blue!60!black]
    (2.0, 0.95) -- (4.0, 0.95)
    node[midway, below=5pt, font=\tiny, blue!60!black] {precision};

  \node[font=\tiny, text width=3.5cm, align=center, gray] at (2.0, 0.1)
    {decoupled: scale and precision\\vary independently};
\end{scope}

\end{tikzpicture}
\caption[Numeric type hierarchy and HDR representation]{%
  Numeric type hierarchy and HDR representation.
  (a)~The precision hierarchy of floating-point types available in
  \FractalShark{}, from native IEEE~754 types (bottom) through
  multi-component expansion types to arbitrary-precision arithmetic (top).
  The dashed outline indicates that any GPU-resident type may be wrapped by
  \code{HDRFloat<T>} to provide wide dynamic range.
  (b)~Comparison of the standard IEEE~754 \code{double} layout, where
  exponent and mantissa share a fixed bit budget, with the
  \code{HDRFloat<T>} representation, which stores an explicit integer
  exponent separately from a fixed-precision mantissa of type~\code{T}.}
\label{fig:type-hierarchy}
\end{figure}


\section{Extended-Precision Floating-Point Types for Perturbation}
\label{sec:extended-precision-types}

In perturbation-based rendering (\cref{sec:perturbation-concept}), each pixel
computes a low-precision \emph{delta} from a high-precision reference orbit.
The precision of this delta arithmetic (the \code{SubType} template parameter)
determines both the accuracy of the rendered image and the GPU throughput.
\FractalShark{} provides a hierarchy of multi-component floating-point types
that trade additional ALU operations for higher effective precision without
relying on native 64-bit floating-point hardware.

\subsection{Motivation: the perturbation precision gap}
\label{subsec:ep-motivation}

Single-precision IEEE~754 (\code{float}, 24-bit mantissa) is the fastest
arithmetic available on consumer GPUs, but it does not always provide enough
precision for accurate perturbation rendering.  In some cases,
rendering artifacts appear because the per-pixel deltas lose significant bits
during the perturbation recurrence.  Native \code{double} (53-bit mantissa)
resolves these artifacts but incurs a substantial throughput penalty on consumer
GPUs, where fp64 performance is often 32:1 slower than fp32.

The \textbf{2$\times$32 double-float} type bridges this gap: it provides
${\sim}48$~bits of effective mantissa using only fp32 arithmetic, delivering
sufficient precision for accurate perturbation while remaining on the fast
fp32 ALUs.  This is the extended-precision type with real practical value in
\FractalShark{}'s rendering pipeline.

\FractalShark{} also includes higher-precision types (2$\times$64,
4$\times$32, 4$\times$64) as reference implementations based on existing work,
primarily for experimentation and completeness.  Some base (non-perturbation)
kernels also use these types for direct Mandelbrot iteration (e.g.\
\code{mandel\_2x\_float}, \code{mandel\_4x\_float}), but those kernels are not
particularly useful in practice---perturbation with linear approximation is the
primary rendering path at any serious zoom depth.

\subsection{The 2\texorpdfstring{$\times$}{x}32 double-float type (\code{dblflt})}
\label{subsec:ep-dblflt}

The core type is \code{MattDblflt} (aliased as \code{dblflt}), a struct
containing two \code{float} values:
\begin{description}
  \item[\code{head}] The most significant bits of the represented value.
  \item[\code{tail}] A correction term satisfying
    $|\mathtt{tail}| \le \tfrac{1}{2}\,\mathrm{ulp}(\mathtt{head})$,
    so that the represented value is exactly $\mathtt{head} + \mathtt{tail}$.
\end{description}
Construction from a \code{double} uses Knuth's error-free sum
algorithm \cite{Knuth1997TAOCP2} to
split the value into a head--tail pair.  The total storage is 8~bytes---the
same footprint as a single \code{double}.

\paragraph{Arithmetic operations.}
The CUDA device functions in \code{dblflt.cuh} implement the standard
operations:

\begin{center}
\begin{tabular}{lll}
\hline
\textbf{Function} & \textbf{Operation} & \textbf{Max relative error} \\
\hline
\code{add\_dblflt(a, b)} & $a + b$  & $O(u^2) \approx 2^{-48}$ \\
\code{sub\_dblflt(a, b)} & $a - b$  & $O(u^2) \approx 2^{-48}$ \\
\code{mul\_dblflt(a, b)} & $a \times b$  & $O(u^2) \approx 2^{-47}$ \\
\code{sqr\_dblflt(a)}    & $a^2$         & optimized via $2 \cdot \mathtt{head} \cdot \mathtt{tail}$ \\
\code{mul\_dblflt2x(a,b)}& $2ab$         & multiply then shift \\
\hline
\end{tabular}
\end{center}

All operations are built on fused multiply-add (\code{\_\_fmaf\_rn},
\code{\_\_fmaf\_rd}) intrinsics, which compute $a \times b + c$ in a single
rounding step.  This is essential for maintaining the head--tail invariant:
intermediate products must be exact or correctly rounded before the correction
term is computed.  The error bounds listed above are the theoretical
$O(u^2)$ bounds for compensated float arithmetic, where $u = 2^{-24}$ is
the float unit roundoff.  Note that the source code comments in
\code{dblflt.cuh} cite empirical error values of ${\sim}2^{-104.7}$;
however, these measurements appear to have been inherited from the
double-double (\code{dbldbl}) implementation (where they are consistent
with $u = 2^{-53}$) and were not re-measured for the float-float variant.

\subsection{\code{CudaDblflt} wrapper and \code{HDRFloat} integration}
\label{subsec:ep-cudadblflt}

The \code{CudaDblflt<T>} class template wraps a \code{dblflt} value with
C++ operator overloads (\code{+}, \code{-}, \code{*}, \code{/}, comparisons)
suitable for use in GPU device code.  This allows perturbation kernels to be
written generically over the \code{SubType} parameter without
\code{dblflt}-specific function calls.

For perturbation rendering with linear approximation, the delta values need
both extended precision \emph{and} wide dynamic range.  This is achieved by
combining \code{CudaDblflt} with the HDR wrapper
(\cref{sec:hdr-float}):
\begingroup\footnotesize
\[
  \code{HDRFloat<CudaDblflt<dblflt>>} \;=\;
    \underbrace{\text{2$\times$32-bit mantissa}}_{\sim\text{48 bits}}
    \;+\;
    \underbrace{\text{int32 exponent}}_{\text{wide dynamic range}}
    \;=\; \text{12 bytes total.}
\]
\endgroup
The \code{Reduce()} method on \code{HDRFloat} normalizes the two float
components relative to the shared exponent by extracting and aligning their
IEEE~754 exponent fields.  A \code{DoubleTo2x32Converter} utility handles
host-side fallback paths where CUDA device code is unavailable.

In the render algorithm enumeration, kernel variants using this type are named
with the \code{2x32} or \code{HDRx2x32} suffix (e.g.\
\code{Gpu2x32PerturbedLAv2}, \code{GpuHDRx2x32PerturbedLAv2}).

\subsection{Higher-precision types (academic interest)}
\label{subsec:ep-higher-prec}

\FractalShark{} includes additional multi-component types based on existing
implementations.  These are primarily of academic interest and are not
essential for typical rendering.

\paragraph{2\texorpdfstring{$\times$}{x}64 double-double (\code{dbldbl}).}
\code{MattDbldbl} pairs two \code{double} values (head + tail) for
${\sim}104$~bits of mantissa in 16~bytes.  On GPUs this is less interesting
than \code{dblflt} because it exercises the slow fp64 units---if fp64
throughput is available, a single \code{double} is usually sufficient for
perturbation, and if more precision is needed, the 4$\times$32 type avoids
fp64 entirely.

\paragraph{4\texorpdfstring{$\times$}{x}32 quad-float (\code{GQF::gqf\_real}).}
Four \code{float} values (${\sim}96$-bit mantissa, ${\sim}29$ decimal digits)
stored as a CUDA \code{float4} vector.  The implementation is adapted from the
QD library by Bailey and Hida \cite{Hida2001QD} (Lawrence Berkeley National Laboratory) and
resides in the \code{GQF} namespace with a \code{make\_qf()} constructor.
Because it uses only fp32 ALUs, it can, in principle, be faster than fp64-based alternatives
on consumer GPUs despite using four components.

\paragraph{4\texorpdfstring{$\times$}{x}64 quad-double (\code{GQD::gqd\_real}).}
Four \code{double} values (${\sim}212$-bit mantissa, ${\sim}64$ decimal
digits) stored as an aligned \code{double4} type.  This is the highest
software-only precision available before switching to arbitrary-precision
arithmetic (MPIR \cite{MPIR_Library} on CPU or \code{HpSharkFloat} on GPU).  Also adapted from
the QD library.

Unlike \code{dblflt}, the quad types do not have a \code{CudaDblflt}-style
wrapper class; they use direct namespaced helper functions for arithmetic.

\subsection{Precision hierarchy}
\label{subsec:ep-hierarchy}

The complete precision hierarchy available for per-pixel computation in
\FractalShark{}, from lowest to highest:

\begin{center}
\begin{tabularx}{\textwidth}{llrX}
\hline
\textbf{Type} & \textbf{Components} & \textbf{Effective bits} & \textbf{Practical role} \\
\hline
\code{float}                & 1$\times$fp32 & ${\sim}24$  & fastest; sometimes insufficient \\
\code{dblflt}               & 2$\times$fp32 & ${\sim}48$  & practical sweet spot \\
\code{double}               & 1$\times$fp64 & ${\sim}53$  & standard; slow on consumer GPUs \\
\code{dbldbl}               & 2$\times$fp64 & ${\sim}104$ & academic \\
\code{GQF::gqf\_real}       & 4$\times$fp32 & ${\sim}96$  & academic \\
\code{GQD::gqd\_real}       & 4$\times$fp64 & ${\sim}212$ & academic \\
MPIR / \code{HpSharkFloat}  & $N$ limbs     & arbitrary   & reference orbit only \\
\hline
\end{tabularx}
\end{center}

Any of these types can be combined with the \code{HDRFloat} exponent wrapper
when wide dynamic range is also required.  In practice, the vast majority of
rendering uses either \code{float}, \code{double}, or
\code{HDRFloat<CudaDblflt<dblflt>>} as the perturbation delta type.

The numeric types introduced here are put to use in the following chapter,
which develops the perturbation framework that makes deep-zoom rendering
practical by reducing per-pixel arithmetic from full precision to lightweight
floating-point deltas.

