\subsection{Cross-Platform Considerations}
\label{sec:cross-platform}

\FractalShark{} is developed primarily as a Windows application built with
Visual~Studio~2026, MSBuild, and the NVIDIA CUDA toolkit. Version~0.53 is the
first release with experimental Linux support through a native target built
with CMake and Clang, with release binaries targeted at Ubuntu. The Linux build
includes the CPU numeric core, CUDA rendering kernels, GPU high-precision
arithmetic, the command-line renderer, the portable unit tests, and the
\code{FractalSharkGuiLinux} graphical application. Windows remains the
historical primary development environment, and the two graphical front ends
do not yet have complete behavioral parity.

\subsubsection{Dual build system}
\label{subsec:dual-build}

The Windows working tree and Visual~Studio solution are authoritative. Shared
projects maintain parallel \code{.vcxproj} and \code{CMakeLists.txt} source
lists and settings. Projects that remain Windows-only do not have CMake targets;
the Win32 GUI and native Linux GUI are intentionally separate applications over
the same renderer and portable command contracts. Adding or removing a source
file from a shared project therefore requires updating both build descriptions.

The preferred Linux build command is \code{./build\_linux.sh}. It configures
and builds persistent \code{build-debug} and \code{build-release} trees. An
incremental build may instead use \code{cmake -{}-build build-debug
-{}-parallel}, with the corresponding release directory used for an
incremental release build. Ordinary C++ translation units are compiled as
C++23 by \code{clang++}. CUDA translation units use C++20 through
\code{nvcc}, with the system \code{g++} selected as the CUDA host compiler.

The Linux target set comprises \code{FractalSharkPlatform},
\code{FractalSharkLib}, \code{FractalSharkGpuLib},
\code{HpSharkFloatLib}, the explicit-instantiation generator, the GPU test
harness and its support library, \code{FractalSharkTest},
\code{FractalSharkCli}, and \code{FractalSharkGuiLinux}. The GUI additionally
requires Xlib, GLX, and OpenGL. Dear~ImGui is fetched at a pinned version during
CMake configuration. Splash images and Imagina test fixtures are embedded into
ELF objects during the build, so the packaged executables do not depend on
loose copies of those assets. The splash-image embedding step currently limits
the GUI target to x86-64 ELF systems.

Continuous integration configures and builds both Debug and Release on Linux,
runs \code{FractalSharkTest}, and smoke-tests a release CLI render. It verifies
that the CLI and GUI binaries have no unresolved shared-library dependencies
and packages both release executables in the Linux archive. The GUI is built
but is not launched on the headless runner. \code{HpSharkFloatTest} is also
built but is not run there because its runtime tests require physical CUDA
hardware. Separate workflow jobs build and test Windows and compile this
document; release publication depends on all three jobs.

\subsubsection{Platform abstraction layer}
\label{subsec:platform-abstraction}

The \code{FractalSharkPlatform} library concentrates operating-system behavior
behind portable interfaces. The \code{Environment} namespace covers aligned
and system-heap allocation, file and directory operations, process and memory
information, timing, thread naming and affinity, console access, input polling,
temporary paths, embedded resources, diagnostics, and small user-interface
helpers. Portable geometry and opaque-handle types prevent Win32 types from
leaking into shared library interfaces.

Dedicated abstractions cover crash handling, process memory limits, wait-cursor
state, and native OpenGL contexts. Each abstraction has separate Win32 and
Linux translation units. The OpenGL wrapper uses WGL on Windows and GLX on
Linux while presenting the same client-rectangle, context, buffer-swap,
renderer-detection, and texture-limit interface to the rendering code. On
Windows, the crash handler writes a \code{.dmp} minidump for post-mortem
analysis.

This separation avoids distributing platform conditionals throughout the
numeric and rendering libraries. Platform-specific graphical behavior remains
in \code{FractalShark::Win32} or \code{FractalShark::Linux}; portable commands
and menu contracts remain directly in \code{FractalShark}. Operating-system
entry points and third-party callback interfaces remain in the global namespace
when required by their APIs.

\subsubsection{Native Linux graphical application}
\label{subsec:linux-gui}

\code{FractalSharkGuiLinux} uses Xlib for its top-level window and event loop,
GLX for the native OpenGL context, and the shared OpenGL display pipeline for
presenting completed frames. The GUI thread owns the window, GLX context, and
all Dear~ImGui state. It forwards input to an in-project ImGui Xlib backend,
presents the latest renderer output, draws overlays, and swaps buffers from the
same event loop. The backend handles mouse buttons and wheels, keyboard input,
focus changes, window resizing, and clipboard integration.

The right-click menu is a native X11 popup hierarchy rather than an ImGui
window. It is generated from the shared \code{MenuTree} catalog and dispatches
through the same portable command identifiers as the Win32 GUI. A shared
\code{MenuState} implementation supplies enablement, check marks, and radio
selections to both native menu builders. Linux-specific command handlers
translate commands that require X11 operations. This arrangement keeps algorithm selection,
navigation, built-in views, iteration controls, perturbation modes, palettes,
tests, benchmarks, and linear-approximation settings synchronized with the
portable command catalog while leaving window-system operations in the Linux
front end.

Dear~ImGui supplies the overlay surfaces that do not need separate native
windows: help and rendering-information messages, file selection, saved-view
selection, manual location entry, and the drag-zoom rectangle. The X11 front
end also implements resize-driven recalculation, mouse-wheel zoom, captured
drag zoom, fullscreen and square-fullscreen transitions, minimization, feature
finding and automatic zoom, save and load commands, and an X11 clipboard owner
for UTF-8 text. A separate Xlib splash window displays a randomly selected
embedded image while initialization proceeds.

The Linux GUI is functional but is not represented as completely equivalent to
the Win32 GUI. Palette rotation currently changes internal state without visibly
recoloring the displayed frame. These limitations should be stated explicitly
rather than hidden behind a general claim of full parity.

\subsubsection{Arbitrary-precision serialization}
\label{subsec:mpir-portability}

\FractalShark{} relies on MPIR for arbitrary-precision integer arithmetic on
Windows and on GMP for the same role on Linux. The two libraries are
source-compatible at the API level but differ in some internal representation
details, which matters because reference-orbit files and saved views contain
serialized arbitrary-precision integers that must round-trip across platforms.

To guarantee byte-identical files, \FractalShark{} defines a single canonical
wire format that matches MPIR's native \code{\_\_mpz\_struct} layout. On
Windows this format is produced by direct field access into the underlying
MPIR object. On Linux the same byte sequence is reconstructed atop GMP using
\code{mpz\_export} and \code{mpz\_import}, which expose limb data through a
portable interface. The two implementations are validated against a checked-in
golden file (\code{golden\_mpir.bin}) by the \code{TestMpirSerialization}
unit test, which runs on both platforms in CI. Any divergence in the wire
format fails the test on whichever platform regressed, preventing silent
incompatibility between Windows-written and Linux-written orbit files.
